---
title: "DroughtNet-NutNet MAPVar analyses"
author: "Tim Ohlert"
date: "11/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##load packages

```{r}
library(tidyverse)
library(plyr)
library(visreg)
library(cowplot)
library(segmented)
```



##create function to calculate D

```{r}
d.fun <- function (var, k=0){
  var<-var[!is.na(var)]
  if(length(var)<3){return(NA)}
  if(sd(var)==0){return(0)}
  if (min(var)<0){var <- var + abs(min(var)) + 0.01*(max(var)-min(var))}
  if (length(var)<2){return(NA)}else{
    var <- var + k
    aux <- numeric(length(var)-1)
    for (i in 1:(length(var)-1)){
      aux [i] <- abs(log(var[i+1]/var[i]))
    }
  }
  return(mean(aux, na.rm=T))}
```


## Load and format climate data for both NutNet and DroughtNet

```{r}
CRU <- read.csv("C:/Users/ohler/Dropbox/Tim Work/NN_DN-CRU-1961-2019.csv")
CRU <- ddply(CRU, .(site_code, year),
             function(x)data.frame(
               ppt = sum(x$ppt)
             ))
CRU <- subset(CRU, year >= "1961" & year <= "2010")

climate_frame <- ddply(CRU, .(site_code),
                       function(x)data.frame(
                         d.var = d.fun(x$ppt),
                         MAP = mean(x$ppt),
                         cv = sd(x$ppt)/mean(x$ppt),
                         sd = sd(x$ppt)
                       ))

```




## Format DroughtNet biomass data

```{r}
reduced_npp <- read.csv("C:\\Users\\ohler\\Dropbox\\IDE MS_Single year extreme\\Data\\anpp_clean_trt_07-23-2020.csv")

temp <- reduced_npp[,c("site_code", "year")]
temp <- unique(temp)
temp <- ddply(temp, c("site_code"),
      function(x)(
        n.year=length(x$year)
      ))
temp$n.year <- temp$V1
reduced_npp <- merge(reduced_npp, temp, by = "site_code", all.x=TRUE)

#only use plots that are not manipulated
control_biomass_DN <- reduced_npp %>% 
  subset( n_treat_days < 30 | trt == "Control")%>%
  subset(n.year > 2)%>%
  ddply( c("site_code", "year", "plot", "subplot"),
         function(x)data.frame(
           biomass = sum(x$mass)
         )) %>%
  #ddply( c("site_code", "year", "plot"),
  #       function(x)data.frame(
   #        biomass = mean(x$biomass)
   #      )) %>%
  ddply( c("site_code", "year"),
         function(x)data.frame(
           biomass = mean(x$biomass)
         )) %>%
  ddply( c("site_code"),
         function(x)data.frame(
           biomass = mean(x$biomass),
           n.year = length(x$year)
         )) %>%
  merge(climate_frame, by = "site_code",all.x=TRUE)%>%
  subset(MAP < 1500)%>%
  subset( site_code != "prades.es")%>%
  subset( site_code != "qdtnorth.cl")

control_biomass_DN <- control_biomass_DN[!is.na(control_biomass_DN$n.year),]
control_biomass_DN$network <- "DroughtNet"

```





## Format NutNet biomass data

```{r}
full.biomass <- read.csv("C:/Users/ohler/Dropbox/NutNet data/full-biomass-24-September-2020.csv")


control_biomass_NN <- full.biomass %>%
  subset(live != 0)%>%
  subset(year_trt == 0 | trt=="Control")%>%
  #block, plot, subplot
  ddply( c("site_code", "year", "plot", "subplot"),
         function(x)data.frame(
           biomass = sum(x$mass)
         )) %>%
  ddply( c("site_code", "year", "plot"),
         function(x)data.frame(
           biomass = mean(x$biomass)
         )) %>%
  ddply( c("site_code", "year"),
         function(x)data.frame(
           biomass = mean(x$biomass)
         )) %>%
  ddply( c("site_code"),
         function(x)data.frame(
           biomass = mean(x$biomass),
           n.year = length(x$year)
         )) %>%
  subset(n.year > 2)%>%
  merge(climate_frame, by = "site_code",all.x=TRUE)%>%
  subset(MAP < 1500)

control_biomass_NN$network <- "NutNet"

```

## Merge DroughtNet and NutNet biomass data

```{r}

dn_nn <- rbind(control_biomass_DN,control_biomass_NN)
dn_nn <- subset(dn_nn, biomass < 2000)
dn_nn <- subset(dn_nn, MAP != "NA")
```




## MAP models

### Look at distribution of MAP values in dataset
```{r}
hist(dn_nn$MAP)
```


### Run the actual models

```{r}
map <- lm(biomass~MAP, data=dn_nn)
map.2 <- lm(biomass~MAP+I(MAP^2), data=dn_nn)
d <- lm(biomass~d.var, data=dn_nn)
map.d <- lm(biomass~MAP * d.var, data=dn_nn)
map.d.2 <- lm(biomass~(MAP+I(MAP^2)) * d.var, data=dn_nn)

```


### Compare AIC scores
```{r}
AIC(map,map.2,d,map.d, map.d.2)
```

### look at model outputs

```{r}
summary(map)
summary(map.2)
summary(d)
summary(map.d)
summary(map.d.2)
```



### Plot MAP model
```{r}
ggplot(dn_nn, aes(x=MAP, y=biomass))+
  geom_smooth(method="lm",formula= (y ~ x + I(x^2)),se=T,col="black",size=2,fill = "dodgerblue", alpha=.2)+
  geom_point(aes(size=d.var, colour=network),shape=19)+
  xlim(c(0,3000))+
  ylab("ANPP (g/m2)")+
  theme_cowplot()
```

### Plot multiple regression
```{r}
visreg2d(map.d, xvar="MAP",yvar="d.var", plot.type="gg", ylab="Variability", main="ANPP")+
  geom_point(data=dn_nn,aes(x=MAP,y=d.var))
```

### MAP model with variability visualized

```{r}
ggplot(dn_nn, aes(x=MAP, y=biomass))+
  geom_smooth(method="lm",formula= (y ~ x),se=T,col="black",size=2,fill = "dodgerblue", alpha=.2)+
  geom_point(aes(size=d.var, colour=network),shape=19)+
  xlim(c(0,1200))+
  ylab("ANPP (g/m2)")+
  theme_cowplot()
```



## Models with residuals

### grab MAP model residuals

```{r}
map.resid <- residuals(map.2)
#map.resid <- resid(map.2)
#x <- c(unname(map.resid))
dn_nn$map.resid <- map.resid
```

### models of residuals

```{r}
mod.resid <- lm(map.resid~d.var,data=dn_nn)
mod.resid.2 <- lm(map.resid~d.var+I(d.var^2),data=dn_nn)
```

### AIC scores and model results

```{r}
AIC(mod.resid,mod.resid.2)
summary(mod.resid.2)
```

### look at results

```{r}
visreg(mod.resid.2)
```


### test for segments

```{r}
pscore.test(mod.resid)
seg <- segmented(mod.resid, seg.Z=~d.var)
```


### plot the segmented

```{r}
plot(seg)  
points(dn_nn$d.var, dn_nn$map.resid)
ylim(-400,900)

```


### plot something else

```{r}
ggplot(lm(biomass~MAP, data=dn_nn)) + 
  geom_point(aes(x=dn_nn$d.var, y=.resid),color = "springgreen4")+
  geom_smooth(aes(x=dn_nn$d.var, y=.resid),method="lm",formula= (y ~ x + I(x^2)),se=T,col="black",size=2,fill = "darkseagreen3")+
  ylab("MAP model residuals")+
  xlab("D (variability)")+
  #geom_smooth(aes(x=dn_nn$d.var, y=.resid),method="lm",se=T,col="black",size=2)+
  theme_bw()
```


### another plot of something

```{r}
plot(dn_nn$d.var, map.resid, xlab="variability (D)",ylab="MAP regression residuals")
abline(0,0)
```







##MAP variability correlation

### models looking for correlation

```{r}
d_map<- lm(d.var~MAP, data=dn_nn)
d_map.2<- lm(d.var~MAP+I(MAP^2), data=dn_nn)
d_map.3<- lm(d.var~MAP+I(MAP^2)+I(MAP^3), data=dn_nn)
```

###compare models and look at them

```{r}
AIC(d_map,d_map.2,d_map.3)
summary(d_map.3)

```

### view correlations

```{r}
visreg(d_map.3, ylab="variability (D)")
```

###another way to view

```{r}
ggplot(dn_nn, aes(x=MAP, y=d.var))+
  geom_smooth(method="lm",formula= (y ~ x + I(x^2)),se=T,col="black",size=2,fill = "dodgerblue", alpha=.2)+
  geom_point(aes(colour=network),shape=19, size=3)+
  xlim(c(0,2500))+
  ylab("Variability")+
  theme_cowplot()

```



##CV - D check

```{r}
cv_d <- lm(d.var~cv, data = dn_nn)
summary(cv_d)
visreg(cv_d, ylab = "D")
```






##Does variability matter in arid sites?

## prep data

```{r}
arid_sites <- subset(dn_nn, MAP<400)
mid_sites <- subset(dn_nn, MAP>400 & MAP<700)
mesic_sites <- subset(dn_nn, MAP>700 &MAP<1000)
super_mesic_sites <- subset(dn_nn, MAP>1000)
```


## models of arid sites

```{r}
map <- lm(biomass~MAP, data=arid_sites)
map.2 <- lm(biomass~MAP+I(MAP^2), data=arid_sites)
d <- lm(biomass~d.var, data=arid_sites)
map.d <- lm(biomass~MAP * d.var, data=arid_sites)
```


## view

```{r}
ggplot(arid_sites, aes(d.var, biomass))+
  geom_smooth(method="lm",formula= (y ~ x),se=T,col="black",size=2,fill = "dodgerblue", alpha=.2)+
  geom_point(aes(colour=network),shape=19, size=3)+
  #xlim(c(0,2500))+
  ylab("ANPP (g/m2)")+
  xlab("variability")+
  ggtitle("1000+ mm")+
  theme_cowplot()
```



## somthing about segmented/breakpoints

```{r}
# have to provide estimates for breakpoints.
# after looking a the data, 
my.seg <- segmented(map, 
                    seg.Z = ~ MAP,
                    psi = list(MAP = c(400, 700)))

# When not providing estimates for the breakpoints "psi = NA" can be used.
# The number of breakpoints that will show up is not defined
#my.seg <- segmented(my.lm, 
#                    seg.Z = ~ DistanceMeters, 
#                    psi = NA)

# display the summary
summary(my.seg)

my.fitted <- fitted(my.seg)

```















